---
title: "Visualizing Numerical Data"
author: "R package build"
date: '2022-02-16'
slug: visualizing-numerical-data
categories: R
tags: GGPLOT2
---

# Loans 

```{r}
library(tidyverse)
library(openintro)
glimpse(loans_full_schema)
```

## Narrowing Down the Data

```{r}
loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)
glimpse(loans)
```

## Histogram 

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram()
```

## Histograms and Experimenting with Binwidth 

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 1000)
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 5000)
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 20000)
```

## Customizing Histograms
```{r}
ggplot(loans, aes(x = loan_amount)) + geom_histogram(binwidth = 5000) +
  labs(x = "Loan amount ($)",y = "Frequency",title = "Amounts of Lending Club loans")
```

## Using Categorical Variables

```{r}
ggplot(loans, aes(x = loan_amount, 
                  fill = homeownership)) +
  geom_histogram(binwidth = 5000,
                 alpha = 0.5) +
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans")
```

## Facet With a Categorical Variable

```{r}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) + 
  geom_histogram(binwidth = 5000) +
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  ) +
  facet_wrap(~ homeownership, nrow = 3)
```

## Density Plot 

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density()
```

## Desity Plots and Adjusting Binwidth
Adjust 0.5
```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 0.5)
```
Adjust 1
```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 1)
```
Adjust 2
```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 2)
```

## Customizing Density Plots 

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 2) +
  labs(
    x = "Loan amount ($)",
    y = "Density",
    title = "Amounts of Lending Club loans"
  )
```

## Adding a Categorical Variable to a Density Plot 
```{r}
ggplot(loans, aes(x = loan_amount, 
                  fill = homeownership)) +
  geom_density(adjust = 2, 
               alpha = 0.5) +
  labs(
    x = "Loan amount ($)",
    y = "Density",
    title = "Amounts of Lending Club loans", 
    fill = "Homeownership"
  )
```

## Boxplots 

```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot()
```

## Boxplots and Outliers

```{r}
ggplot(loans, aes(x = annual_income)) +
  geom_boxplot()
```

## Customizing Box Plots 

```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot() +
  labs(
    x = "Interest rate (%)",
    y = NULL,
    title = "Interest rates of Lending Club loans"
  ) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

## Adding a Categorical Variable to a Box Plot

```{r}
ggplot(loans, aes(x = interest_rate,
                  y = grade)) +
  geom_boxplot() +
  labs(
    x = "Interest rate (%)",
    y = "Grade",
    title = "Interest rates of Lending Club loans",
    subtitle = "by grade of loan"
  )
```

## Scatter Plots

```{r}
ggplot(loans, aes(x = debt_to_income, y = interest_rate)) +
  geom_point()
```



