---
title: "Importing Data"
author: "R package build"
date: '2022-02-22'
slug: importing-data
categories: R
tags: []
---

# Reading rectangular data into R

### Readr
+ read_csv() - comma delimited files
+ read_csv2() - semicolon separated files (common in countries where , is used as the decimal place)
+ read_tsv() - tab delimited files
+ read_delim() - reads in files with any delimiter
+ read_fwf() - fixed width files

### Readxl 
+ read_excel() - read xls or xlsx files

## Reading Data
```{r, results=FALSE}
library(tidyverse)
url <- 'https://raw.githubusercontent.com/angelacalicchio/tidyverse/main/nobel.csv'
nobel <- read.csv(url)
nobel
```

## Writing Data

### Write a File
```{r}
url <- 'https://raw.githubusercontent.com/angelacalicchio/tidyverse/main/df-na.csv'
df_na <- read.csv(url)
```

```{r}
df <- tribble(
  ~x, ~y,
  1,  "a",
  2,  "b",
  3,  "c"
)
write_csv(df, file = "df_na")
read_csv("df_na")
```

## Data with Bad Names
```{r}
url <- 'https://raw.githubusercontent.com/angelacalicchio/tidyverse/main/edibnb-badnames.csv'
edibnb_badnames <- read.csv(url)
names(edibnb_badnames)
```

...but R doesn't allow spaces in variable names

```{r, error=TRUE}
ggplot(edibnb_badnames, aes(x = Number of bathrooms, y = Price)) +
  geom_point()
```

> We have to find a way to get by this error

### Option 1: Define Column Names 
```{r}
edibnb_col_names <- read_csv(url,
                             col_names = c("id", "price", 
                                           "neighbourhood", "accommodates",
                                           "bathroom", "bedroom", 
                                           "bed", "review_scores_rating", 
                                           "n_reviews", "url"))
names(edibnb_col_names)
```

### Option 2: Format Text to snake_case
```{r}
edibnb_clean_names <- read_csv(url) %>%
  janitor::clean_names()
names(edibnb_clean_names)
```

## Variable Types
```{r, results=FALSE}
read.csv(url)
```
### Option 1: Explicit NAs
```{r}
url <- 'https://raw.githubusercontent.com/angelacalicchio/tidyverse/main/df-na.csv'
df_na <- read.csv(url)
read_csv(url, 
         na = c("", "NA", ".", "9999", "Not applicable"))
```

### Option 2: Specify Column Types
```{r}
read_csv(url, col_types = list(col_double(), 
                                            col_character(), 
                                            col_character()))
```

## Favorite Foods
```{r}
url <- 'https://raw.githubusercontent.com/angelacalicchio/tidyverse/main/favourite-food.csv'
fav_food <- read.csv(url)
fav_food
```

### Variable Names
```{r}
fav_food <- read.csv(url) %>%
  janitor::clean_names()
fav_food
```

### Handling NAs
```{r}
fav_food <- read.csv(url,
                       na = c("N/A", "99999")) %>%
  janitor::clean_names()
fav_food
```

### Make **Age** Numeric
```{r}
fav_food <- fav_food %>%
  mutate(
    age = if_else(age == "five", "5", age),
    age = as.numeric(age)
    )
glimpse(fav_food)
```

### Socio-economic Status
```{r}
fav_food %>%
  count(ses)
```

### Make **ses** Factor
```{r}
fav_food <- fav_food %>%
  mutate(ses = fct_relevel(ses, "Low", "Middle", "High"))
fav_food %>%
  count(ses)
```

## Putting it all together
```{r}
fav_food <- read.csv(url, na = c("N/A", "99999")) %>%
  janitor::clean_names() %>%
  mutate(
    age = if_else(age == "five", "5", age), 
    age = as.numeric(age),
    ses = fct_relevel(ses, "Low", "Middle", "High")
  )
fav_food
```

## Out and Back in 
```{r}
url_2 <- 'https://raw.githubusercontent.com/angelacalicchio/tidyverse/main/fav-food-clean.csv'
fav_foods_clean <- read.csv(url_2)
fav_foods_clean
```
SES
```{r}
fav_foods_clean %>%
  count(ses)
```

## read_rds() and write_rds()
+ CSVs can be unreliable for saving interim results if there is specific variable type information you want to hold on to.
+ An alternative is RDS files, you can read and write them with read_rds() and write_rds(), respectively.
```{r, eval=FALSE}
read_rds(path)
write_rds(x, path)
```

