---
title: "Data Classes"
author: "R package build"
date: '2022-02-23'
slug: data-classes
categories: R
tags: []
---

# Data Classes

```{r}
library(tidyverse)
url <- 'https://raw.githubusercontent.com/angelacalicchio/tidyverse/main/cat-lovers.csv'
cat_lovers <- read.csv(url)
```

### Factors
R uses factors to handle categorical variables, variables that have a fixed and known set of possible values
```{r}
x <- factor(c("BS", "MS", "PhD", "MS"))
x
```
```{r}
typeof(x)
class(x)
```
We can think of factors like character (level labels) and an integer (level numbers) glued together
```{r}
glimpse(x)
```
```{r}
as.integer(x)
```

### Dates
```{r}
y <- as.Date("2020-01-01")
y
```
```{r}
typeof(y)
class(y)
```
We can think of dates like an integer (the number of days since the origin, 1 Jan 1970) and an integer (the origin) glued together
```{r}
as.integer(y)
```
```{r}
as.integer(y) / 365 # Roughly 50 years
```

### Data Frames
We can think of data frames like like vectors of equal length glued together
```{r}
df <- data.frame(x = 1:2, y = 3:4)
df
```
```{r}
typeof(df)
class(df)
```

### Lists 
```{r}
l <- list(
  x = 1:4,
  y = c("hi", "hello", "jello"),
  z = c(TRUE, FALSE))
l
```

#### Lists and Data Frames 
```{r}
df
```
```{r}
df %>%
  pull(y)
```

## Cat Lovers
```{r}
glimpse(cat_lovers)
```

### Data Coercion when Plotting
First, let's see what the data looks like without **Coercion**
```{r}
ggplot(cat_lovers, mapping = aes(x = handedness)) +
  geom_bar()
```

Now, let's see what the data looks like after using **mutate()** 
```{r}
cat_lovers %>%
  mutate(handedness = fct_infreq(handedness)) %>%
  ggplot(mapping = aes(x = handedness)) +
  geom_bar()
```

## Hotels
```{r}
library(tidyverse)
url <- 'https://raw.githubusercontent.com/angelacalicchio/tidyverse/main/hotels.csv'
hotels <- read.csv(url)
```
### Calculate and visualise the number of bookings on any given arrival date
```{r, results=FALSE}
hotels %>%
  select(starts_with("arrival_"))
```

### Step 1: Contruct Dates
```{r, results=FALSE}
library(glue)
hotels %>%
  mutate(
    arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")
    ) %>% 
  relocate(arrival_date)
```

### Step 2: Count Bookings per Date
```{r, results=FALSE}
hotels %>%
  mutate(arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")) %>%
  count(arrival_date)
```

### Step 3: Visualize Bookings per Date
```{r}
hotels %>%
  mutate(arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")) %>%
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_line()
```

> We need to revise the data and construct dates "as dates"

### Step 1: Revised
```{r, results=FALSE}
library(lubridate)
hotels %>%
  mutate(
    arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))
    ) %>% 
  relocate(arrival_date)
```

### Step 2: Revised
```{r, results=FALSE}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date)
```

### Step 3A: Visualize bookings per date
```{r}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_line()
```

Step 3B: Visualize using a smooth curve
```{r}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_smooth()
```