---
title: "Exploratory Data Analysis"
author: "R package build"
date: '2022-03-26'
slug: exploratory-data-analysis
categories: []
tags: []
---

# Loading and Preparing the Data 
```{r}
# Install and Load Packages
if(!require("pacman")) install.packages("pacman")

pacman::p_load(pacman, magrittr, productplots, psych, RColorBrewer, tidyverse)
```

## "happy" dataset from productplots package
```{r}
?happy
names(happy)
```

## Save dataset to tibble named "df" (for "dataframe")
```{r}
df <- happy %>%
  as_tibble() %>%
  print()
```

## Delete id and wtssall, which is a weighting variable that doesn't change results appreciably 
```{r}
df %<>%
  select(happy:health) %>%
  print()
```

# Check levels of "happy"
```{r}
levels(df$happy)
```

## Reverse level of "happy" so "very happy" is at the top of stacked bar charts 
```{r}
df %<>%
  mutate(happy = fct_rev(happy))

levels(df$happy)
```

# Visual Exploration 

## Bar Charts of Happy 
```{r}
df %>%
  ggplot() + 
  geom_bar(aes(happy, fill = happy)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none")
```

### Frequencies of Happy
```{r}
df %>% count(happy)
```

## Filter out the NA responses on happy
```{r}
df %<>%
  filter(!is.na(happy))

df %>%
  ggplot() + 
  geom_bar(aes(happy, fill = happy)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none")
```

### Frequencies for happy 
```{r}
df %>% count(happy)
```

# Happiness and Gender 

## Bar chart of sex
```{r}
df %>%
  ggplot() + 
  geom_bar(aes(sex, fill = sex)) + 
  theme(
    axis.title.x = element_blank(), 
    legend.position = "none")
```

### Frequencies for sex
```{r}
df %>% count(sex)
```

### 100% stacked bar chart
```{r}
df %>%
  ggplot(aes(sex, fill = happy)) + 
  geom_bar(position = "fill")
```

# Happiness and Marital Status

## Bar chart of marital status
```{r}
df %>%
  ggplot() + 
  geom_bar(aes(marital, fill = marital)) + 
  theme(
    axis.title.x = element_blank(), 
    legend.position = "none")
```

### Frequencies for Marital
```{r}
df %>% count(marital)
```

## 100% stacked bar chart
```{r}
df %>%
  filter(!is.na(marital)) %>%
  ggplot(aes(marital, fill = happy)) + 
  geom_bar(position = "fill")
```

### Create dichotmous married/not married variable
```{r}
df %<>%
  mutate(
    married = ifelse(
      marital == "married",
      "yes",
      "no")
  ) %>%
  mutate(married = as.factor(married)) %>%
 print()
```

### Frequencies for married 
```{r}
df %>% count(married)
```

## 100% stacked bar chart
```{r}
df %>%
  filter(!is.na(married)) %>%
  ggplot(aes(married, fill = happy)) + 
  geom_bar(position = "fill")
```

# Happiness and Level of Education

## Bar chart of degree
```{r}
df %>%
  ggplot() + 
  geom_bar(aes(degree, fill = degree)) + 
  theme(
    axis.title.x = element_blank(), 
    legend.position = "none")
```

### Frequencies of degree
```{r}
df %>% count(degree)
```

## 100% stacked bar chart
```{r}
df %>%
  filter(!is.na(degree)) %>%
  ggplot(aes(degree, fill = happy)) + 
  geom_bar(position = "fill")
```

### Create dichotomous college/not variable for exploration
```{r}
df %<>%
  mutate(
    college = ifelse(
      degree %in%
      c("junior college",
        "bachelor",
        "graduate"),
      "yes", "no")
  ) %>%
  print()
```

### Frequencies of college
```{r}
df %>% count(college)
```

## 100% stacked bar chart
```{r}
df %>%
  ggplot(aes(college, fill = happy)) + 
  geom_bar(position = "fill")
```

# Happiness and Financial Status

## Bar chart of finrela
```{r}
df %>%
  ggplot() + 
  geom_bar(aes(finrela, fill = finrela)) + 
  theme(
    axis.title.x = element_blank(), 
    legend.position = "none")
```

### Frequencies for finrela
```{r}
df %>% count(finrela)
```

## 100% stacked bar chart
```{r}
df %>%
  filter(!is.na(finrela)) %>%
  ggplot(aes(finrela, fill = happy)) + 
  geom_bar(position = "fill")
```

### Create dichotomous varible for average finances or higher
```{r}
df %<>%
  mutate(
    avg_fin = case_when(
      finrela %in%
        c("far below average",
          "below average") ~
          "no",
      finrela %in%
        c("average",
          "above average",
          "far above average") ~
          "yes",
      finrela == "NA" ~ "NA")
  ) %>%
  print()
```

### Get frequencies
```{r}
df %>% count(avg_fin)
```

## 100% stacked bar chart
```{r}
df %>%
  filter(!is.na(avg_fin)) %>%
  ggplot(aes(avg_fin, fill = happy)) + 
  geom_bar(position = "fill")
```

# Happiness and Health

## Bar chart of health
```{r}
df %>%
  ggplot() + 
  geom_bar(aes(health, fill = health)) + 
  theme(
    axis.title.x = element_blank(), 
    legend.position = "none")
```

### Frequencies of health
```{r}
df %>% count(health)
```

## 100% stacked bar chart
```{r}
df %>%
  filter(!is.na(health)) %>%
  ggplot(aes(health, fill = happy)) + 
  geom_bar(position = "fill")
```

# Happiness and Year of Survey

## Histogram of year
```{r}
df %>% qplot(year, binwidth = 5, data = .)
```

### Descriptive statistics for year
```{r}
df %>% select(year) %>% summary()
``` 

## Density plots of year by group
```{r}
df%>%
  ggplot(aes(x = year, 
    fill = happy)) + 
  geom_density(alpha = 0.5) +
  facet_grid(happy ~ .) +  # facet_grid
  theme(legend.position = "none") 
```
## Boxplots of year by group
```{r}
df %>%
  ggplot(aes(x = happy, 
    y = year, 
    fill = happy)) + 
  geom_boxplot() +
  coord_flip() +
  xlab("") +
  theme(legend.position = "none")
```

# Happiness and Age

## Histogram of age
```{r}
df %>% qplot(age, binwidth = 5, data = .)
```

### Descriptive statistics for age
```{r}
df %>% select(age) %>% summary()
```

## Density plots of age by group
```{r}
df %>%
  ggplot(aes(x = age, 
    fill = happy)) + 
  geom_density(alpha = 0.5) +
  facet_grid(happy ~ .) +  # facet_grid
  theme(legend.position = "none")  
```

# Boxplots of age by group
```{r}
df %>%
  ggplot(aes(x = happy, 
    y = age, 
    fill = happy)) + 
  geom_boxplot() +
  coord_flip() +
  xlab("") +
  theme(legend.position = "none")
```

# Happiness and Year Born

### Calculate year of birth
```{r}
df %<>%
  mutate(born = year - age)
```

## Histogram of born
```{r}
df %>% qplot(born, binwidth = 5, data = .)
```

### Descriptive statistics for born
```{r}
df %>% select(born) %>% summary()
```

## Density plots of born by group
```{r}
df %>%
  ggplot(aes(x = born, 
    fill = happy)) + 
  geom_density(alpha = 0.5) +
  facet_grid(happy ~ .) +  # facet_grid
  theme(legend.position = "none") 
```
## Boxplots of born by group
```{r}
df %>%
  ggplot(aes(x = happy, 
    y = born, 
    fill = happy)) + 
  geom_boxplot() +
  coord_flip() +
  xlab("") +
  theme(legend.position = "none")
```

# Clean Up

## Clear data
```{r} 
rm(list = ls()) 
```

## Clear packages
```{r}
detach("package:datasets", unload = T) 

p_unload(all)  # Remove all contributed packages
```

## Clear plots
```{r}
graphics.off()  
```

## Clear console
```{r}
cat("\014")  # Mimics ctrl+L
```
