---
title: 'A02: Regression Models'
author: "R package build"
date: '2022-03-26'
slug: a02-regression-models
categories: []
tags: []
---

# Link to Git Repository Excel File and CSV file 
```{r, link_to_github}
link <- url("https://github.com/angelacalicchio/a02-diamonds/tree/main")
```

# Load and Prepare Data and Packages
```{r, setup}
pacman::p_load(pacman, magrittr, productplots, psych, RColorBrewer, tidyverse)

url <- 'https://raw.githubusercontent.com/angelacalicchio/a02-diamonds/main/Professor_Proposes_Data.csv'
diamond_selection <- read_csv(url)

View(diamond_selection)
```

> **IMPORTANT**
> When viewing the excel file that was uploaded to rstudio, I noticed that price was being percieved as a date. In order to combat this, I converted to excel file to a csv file and uploaded that data to github, then used the raw data to upload it back to rstudio. (Both the excel file and csv file can be viewed from the link above)

# Save dataset to tibble named "df" (for "dataframe")
```{r}
df <- diamond_selection %>%
  as_tibble() %>%
  print()
```

# Data Visualization 

## Categorical Variables 

### Color

#### Creating a Bar Chart looking at Color
```{r}
ggplot(diamond_selection) +
  geom_bar(aes(x = Colour, fill = Colour)) +
  scale_fill_manual(values = c("D" = "#F8766D",
                               "E" = "#F8766D",
                               "F" = "#F8766D",
                               "G" = "#00BFC4",
                               "H" = "#00BFC4",
                               "I" = "#00BFC4",
                               "J" = "#C77CFF",
                               "K" = "#C77CFF",
                               "L" = "#C77CFF",
                               "NA" = "black"))
```

#### Removing NAs from Colour
```{r}
df %<>%
  filter(!is.na(Colour))

df %>%
  ggplot() + 
  geom_bar(aes(Colour, fill = Colour)) +
  scale_fill_manual(values = c("D" = "#F8766D",
                               "E" = "#F8766D",
                               "F" = "#F8766D",
                               "G" = "#00BFC4",
                               "H" = "#00BFC4",
                               "I" = "#00BFC4",
                               "J" = "#C77CFF",
                               "K" = "#C77CFF",
                               "L" = "#C77CFF",
                               "NA" = "black")) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none")
```

#### Frequencies for Color
```{r}
df %>% count(Colour)
```

### Clarity 

#### Creating a Bar Chart looking at Clarity
```{r}
ggplot(diamond_selection) +
  geom_bar(aes(x = Clarity, fill = Clarity)) +
  scale_fill_manual(values = c("I1" = "#F8766D",
                               "I2" = "#F8766D",
                               "SI1" = "#7CAE00",
                               "SI2" = "#7CAE00",
                               "SI3" = "#7CAE00",
                               "VS1" = "#00BFC4",
                               "VS2" = "#00BFC4",
                               "VVS1" = "#C77CFF",
                               "VVS2" = "#C77CFF",
                               "NA" = "black"))
```

#### Removing NAs from Clarity 
```{r}
df %<>%
  filter(!is.na(Clarity))

df %>%
  ggplot() + 
  geom_bar(aes(Clarity, fill = Clarity)) +
  scale_fill_manual(values = c("I1" = "#F8766D",
                               "I2" = "#F8766D",
                               "SI1" = "#7CAE00",
                               "SI2" = "#7CAE00",
                               "SI3" = "#7CAE00",
                               "VS1" = "#00BFC4",
                               "VS2" = "#00BFC4",
                               "VVS1" = "#C77CFF",
                               "VVS2" = "#C77CFF",
                               "NA" = "black")) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none")
```

#### Frequencies for Clarity
```{r}
df %>% count(Clarity)
```

### Cut 

#### Creating a Bar Chart looking at Cut
```{r}
ggplot(diamond_selection) +
  geom_bar(aes(x = Cut, fill = Cut))
```

#### Removing NAs from Cut
```{r}
df %<>%
  filter(!is.na(Clarity))

df %>%
  ggplot() + 
  geom_bar(aes(Clarity, fill = Clarity)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none")
```

#### Frequencies in Cut
```{r}
df %>% count(Cut)
```

### Certification 

#### Creating a Bar Chart looking at Certification
```{r}
ggplot(diamond_selection) +
  geom_bar(aes(x = Certification, fill = Certification))
```

#### Removing NAs
```{r}
df %<>%
  filter(!is.na(Certification))

df %>%
  ggplot() + 
  geom_bar(aes(Certification, fill = Certification)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none")
```

#### Frequencies in Certification
```{r}
df %>% count(Certification)
```

### Polish 

#### Creating a Bar Chart looking at Polish
```{r}
ggplot(diamond_selection) +
  geom_bar(aes(x = Polish, fill = Polish))
```

#### Removing NAs 
```{r}
df %<>%
  filter(!is.na(Polish))

df %>%
  ggplot() + 
  geom_bar(aes(Polish, fill = Polish)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none")
```

#### Frequencies in Polish
```{r}
df %>% count(Polish)
```

### Symmetry 

#### Creating a Bar Chart looking at Symmetry
```{r}
ggplot(diamond_selection) +
  geom_bar(aes(x = Symmetry, fill = Symmetry))
```

#### Removing NAs
```{r}
df %<>%
  filter(!is.na(Symmetry))

df %>%
  ggplot() + 
  geom_bar(aes(Symmetry, fill = Symmetry)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none")
```

#### Frequencies in Symmetry
```{r}
df %>% count(Symmetry)
```

## Continuous Variables 

### Carat

#### Creating a Histogram for Carat
```{r}
ggplot(diamond_selection) +
  geom_histogram(aes(x = Carat), binwidth = 0.1)
```

# Bootstrapping the data

## Preparing the data for bootstrapping
```{r}
library(boot)
library(patchwork)
library(tidyverse)
library(pacman)
```

```{r}
summary(df)
```

## Comparison of Variables with Price, Boxplots
```{r}
carat.plot <- ggplot(data = df, 
       mapping = aes(x = Carat, y = Price, colour = Wholesaler)) +
  geom_point()
  
color.plot <- ggplot (df, aes(x = Colour, y = Price)) +
  geom_boxplot()

cut.plot <- ggplot (df, aes(x = Cut, y = Price)) +
  geom_boxplot()

clarity.plot <- ggplot (df, aes(x = Clarity, y = Price)) +
  geom_boxplot()

certification.plot <- ggplot (df, aes(x = Certification, y = Price)) +
  geom_boxplot()

polish.plot <- ggplot (df, aes(x = Polish, y = Price)) +
  geom_boxplot()

symmetry.plot <- ggplot (df, aes(x = Symmetry, y = Price)) +
  geom_boxplot()
 carat.plot + cut.plot + color.plot + clarity.plot + certification.plot + polish.plot + symmetry.plot 
```
 
> Looking at the graphs, we can determine that Carat and Color have the most impact on price. There seems to be a positive correlation between price and carat, as well as a correlation between certain colors and price. The other factors have lots of variation in them, which may indicate that there is no correlation between them and price or perhaps a small positive correlation. Because of this, the factors that I will focus on from here on out will be carat and color. If we are inferring from the data, the diamond that the Professor is searching for should, in theory, be cheaper than $3000, which is well within his range. Also, if look into the Wholesaler of choice, we can immediately remove wholesaler 3 who has little to no inventory in the area that the Professor is searching for.

## Effect of Variables on the Carat
```{r}
cut <- ggplot (df, aes(x = Cut, y = Carat)) +
  geom_boxplot()
color <- ggplot (df, aes(x = Colour, y = Carat)) +
  geom_boxplot()
clarity <- ggplot (df, aes(x = Clarity, y = Carat)) +
  geom_boxplot()
certification <- ggplot (df, aes(x = Certification, y = Carat)) +
  geom_boxplot()
polish <- ggplot (df, aes(x = Polish, y = Carat)) +
  geom_boxplot()
symmetry <- ggplot (df, aes(x = Symmetry, y = Carat)) +
  geom_boxplot()
cut +  color + clarity + certification + polish + symmetry 
```

> This is very important, because our data shows that carat has a high impact on the price of the diamond, as does color, which was to be expected. By looking at the other variables impacts on carat, we can determine that they have little impact on the price, as they have little impact on carat.

## Creating a Linear Model
```{r}
set.seed(1)
head(df)
dim(df)
## Create an index to randomly sample 50% records from Auto
train <- sample(440, 220)
```
```{r}
lm.fit <- lm(Carat~Price, data = df) 
lm.fit
```
```{r}
plot(lm.fit)
```

## Bootstrapping the Data

### Set Seed
```{r}
set.seed(1)
head(df)
dim(df)
train <- sample(440, 220)
head(train)
```

### Filter Out Wholesaler 3
```{r}
df.1.2 <- df %>%
  filter(Wholesaler == 3)

df.1.2
```

### Graph
```{r}
df %>%
  ggplot(aes(x = Carat, y = Price, color = Wholesaler)) + 
  geom_point()
```

```{r}
df.3 <- df %>%
  filter(Wholesaler == 2 | Wholesaler == 1)
dim(df.3)
df.3 %>% 
  ggplot(aes(x = Carat, y = Price, color = Wholesaler 
    )) +
  geom_point()
attach(df.3)
```

```{r}
dim(df.3)
```

```{r}
boot.fn <- function(data, index) {
  return(coef(lm(Carat~Price + Colour + Clarity+ Cut + Certification + Polish + Symmetry, data = df , subset = train)))
}
set.seed(1)
boot.fn(df, sample(240, 120, replace=T))
```

> The professor found a diamond that seemed to be well within the parimeters of diamonds that are comparable, by my calcualtions a diamond of this caliber should cost around $2900-$3000; however, since there is no diamond within the data set with the exact parimeters, it is safe to assume that the price he is paying is completely normal and he shouldn't have to worry about being ripped off from Wholesaler 1 or 2. 

