---
title: 'Professor Evaluations '
author: "R package build"
date: '2022-02-28'
slug: professor-evaluations
categories: []
tags: []
---

# Preliminaries

## Importing Libraries
```{r}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
```

# Moderndive Case Study

## Importing Data
```{r}
evals_ch5 <- evals %>%
  select(ID, score, bty_avg, age)
```

## Explore the Data
```{r}
glimpse(evals_ch5)
```

```{r}
evals_ch5 %>%
  sample_n(size = 5)
```

## Mean and Median of BTY and Score
```{r}
evals_ch5 %>%
  summarise(mean_bty_average = mean(bty_avg), mean_score = mean(score), median_bty_avg = median(bty_avg), median_score = median(score))
```

## Five Number Summary
```{r}
evals_ch5 %>%
  select(score, bty_avg) %>%
  skim()
```

## Scatter Plot of Score vs. Bty_avg
```{r}
plot(evals_ch5$score, evals_ch5$bty_avg)
```

## Correlation Between Score and Bty_avg
There are two ways to get the correlation 

### Version 1
```{r}
evals_ch5 %>%
  get_correlation(formula = score ~ bty_avg)
```

### Version 2
```{r}
evals_ch5 %>%
  summarise(correlation = cor(score, bty_avg))
```

## Scatter Plot: Score vs Bty_avg
```{r}
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score",
       y = "Teaching Score",
       title = "Scatterplot of Relationship of Teaching and Beauty Scores")
```

## Scatter Plot with geom_jitter()
```{r}
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  labs(x = "Beauty Score",
       y = "Teaching Score",
       title = "Scatterplot of Relationship of Teaching and Beauty Scores")
```

## Linear Regression line on Scatter Plot
```{r}
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score",
       y = "Teaching Score",
       title = "Scatterplot of Relationship of Teaching and Beauty Scores") +
  geom_smooth(method = "lm", se = FALSE)
```

## Linear Regression
```{r}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch5)

# Get regression table:
get_regression_table(score_model)
```

## Fitting the Regression Model
```{r}
regression_points <- get_regression_points(score_model)
regression_points
```
